---
title: "p8105_hw1_kpr2118.Rmd"
author: "Kailey Rishovd"
date: "9/19/2020"
output: github_document
---
# PROBLEM 1 

The chunk below creates a dataframe containing a random sample of size 10 from a std normal distribution, a logical vector indicating if elements are GT 0, a character vector length 10, and a factor vector length 10 with 3 factor levels. 

note: running the tidyverse library in order to use the tibble fxn. 

```{r problem_1}
library(tidyverse)

prob1_df = tibble(
  rand_samp = rnorm(10),
  rand_samp_pos = rand_samp > 0,
  vec_char = c("This", "is", "a", "character", "vector", "of", "length", "ten", "characters", "right"),
  vec_factor = factor(c("first", "second", "third", "first", "second", "third", "first", "second", "third", "first"))
)

prob1_df
```

The code chunk attempts to find the mean of each variable. 

```{r Variable_Means, message=FALSE, warning=FALSE}
mean_rand_samp = mean(pull(prob1_df, rand_samp))
mean_rand_samp
mean_rand_samp_pos = mean(pull(prob1_df, rand_samp_pos))
mean_rand_samp_pos
mean_vec_char = mean(pull(prob1_df, vec_char))
mean_vec_char
mean_vec_factor = mean(pull(prob1_df, vec_factor))
mean_vec_factor

#What works and what doesn't?
#The numeric variables (rand_samp and rand_samp_pos) create means but the non-numeric variables (vec_char and vec_factor) return NA.
```

This chunk of code converts variable type. Character --> numeric. 

```{r Variable_type_convert, eval=FALSE}
as.numeric(pull(prob1_df, rand_samp_pos))
as.numeric(pull(prob1_df, vec_char))
as.numeric(pull(prob1_df, vec_factor))

#What happens and why? 
#Values turn into 1, 0 for rand_samp_pos -- values turn into 1, 2, 3 for vec_factor -- values for vec_char result in NA. Vec_char is non-numeric and has no numeric conversion. This is just like the means example above. 
```

This code chunk converts the logical vector to numeric... 

```{r vector_conversion_part_2}
as.numeric(pull(prob1_df, rand_samp_pos)) * pull(prob1_df, rand_samp_pos)
as.factor(pull(prob1_df, rand_samp_pos)) * pull(prob1_df, rand_samp_pos)
as.numeric(as.factor(pull(prob1_df, rand_samp_pos))) * pull(prob1_df, rand_samp_pos)
```


